---
title: "Data Wrangling"
author: "Sean Li and Ben Thorpe"
date: "TBD"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(tidyverse)
library(hoopR)
```


# Data Wrangling

## Loading in Data

- loading in from packages

```{r}
tictoc::tic()
progressr::with_progress({
  nba_pbp <- hoopR::load_nba_pbp(2016:2021)
})
tictoc::toc()
```

```{r}
glimpse(nba_pbp)
```


- loading in from .csv

- loading in from .xlsx

## Exploring dplyr

dplyr is a package within the tidyverse

- Select,filter, arrange, mutate, summarise
- split into row and column functions 
- splice() and rename() and drop NA
- Have them try a few exercises
- Group_by + summarise 

The "pipe" operator takes the object on the left-hand-side and passes it into
the object on the right-hand-side. Rather than write `f(x, y)` one could
write `x %>% f(y)`.

Let's `select()` the `shooting_play`, runs allowed, wins, year, and team name.

```{r}
# teams %>% 
#  select(R, RA, W, yearID, name)
```



## Try it yourself 
- have a few simple exercises to practice basic dplyr functions
- maybe one of each function


## Combining it together

- cover join stuff
- getting min and max of some variables




