---
title: "Data Wrangling"
author: "Sean Li and Ben Thorpe"
date: "TBD"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(tidyverse)
library(hoopR)
library(openxlsx)
```


# Data Wrangling

## Loading in Data

- loading in from packages

```{r}
tictoc::tic()
progressr::with_progress({
  nba_pbp <- hoopR::load_nba_pbp(2021)
})
tictoc::toc()
```

```{r}
glimpse(nba_pbp)
```

- loading in from .csv

```{r}
#csv_data <- read.csv("data/2019-20_pbp.csv")
```


- loading in from .xlsx

```{r}
xlsx_data <- read.xlsx(xlsxFile = "data/nbadata.xlsx")
```

## Exploring dplyr

dplyr is a package within the tidyverse

- Select,filter, arrange, mutate, summarise
- split into row and column functions 
- splice() and rename() and drop NA
- Have them try a few exercises
- Group_by + summarise 

The "pipe" operator takes the object on the left-hand-side and passes it into
the object on the right-hand-side. Rather than write `f(x, y)` one could
write `x %>% f(y)`.

Let's `select()` the `period_number`and `clock_display_value`.

```{r}
clock_data <- nba_pbp %>% 
  select(period_number, clock_display_value) %>%
  head(10000)
clock_data
```

Filters:

```{r}
nba_pbp %>%
  filter(home_score == 100 & away_score == 100)

nba_pbp %>%
  filter((home_score > 100 & home_team_mascot == "Heat") | (away_score > 100 & away_team_mascot == "Heat")) %>%
  select(home_score, away_score, home_team_mascot, away_team_mascot)

nba_pbp %>%
  filter(home_score == c(120, 130, 140)) %>%
  arrange(desc(home_score)) %>%
  select(period_display_value, home_team_abbrev, away_team_abbrev, home_score, away_score)

nba_pbp %>%
  filter(grepl('LeBron', text)) %>%
  head(1) %>%
  select(text) %>%
  pull()
```


Quick groupby intro:

```{r}
nba_pbp %>%
  group_by(type_text) %>%
  count() %>%
  arrange(desc(n)) %>%
  select(type_text) %>%
  head(5)
  

nba_pbp %>%
  group_by(type_text) %>%
  summarise(mean_secs = mean(start_quarter_seconds_remaining), sd_secs = sd(start_quarter_seconds_remaining))
```

Dropping NA values:

```{r}
nba_pbp %>% drop_na()

nba_pbp %>% drop_na(participants_1_athlete_id)
```

Creating new columns with `mutate()`:

```{r}
nba_pbp %>%
  mutate(total_score = home_score + away_score) %>%
  select(home_score, away_score, total_score)

nba_pbp %>%
  mutate(home_team_favored = ifelse(game_spread < 0, TRUE, FALSE)) %>%
  select(game_spread, home_team_favored, home_team_abbrev) %>%
  arrange(desc(home_team_favored))
```


## Example: Who had the best record against the spread?

```{r}
spread_data <- nba_pbp %>%
  group_by(game_id) %>%
  slice(which.max(game_play_number)) %>%
  select(game_spread, home_team_abbrev, away_team_abbrev, home_score, away_score) %>%
  mutate(score_diff = away_score - home_score, home_covered = ifelse(score_diff<game_spread, TRUE, FALSE),
         away_covered = ifelse(score_diff>game_spread, TRUE, FALSE))
spread_data

spread_home <- spread_data %>% 
  group_by(home_team_abbrev, home_covered) %>%
  count() %>%
  pivot_wider(names_from = home_covered, values_from = n)
spread_home

spread_away <- spread_data %>% 
  group_by(away_team_abbrev, away_covered) %>%
  count() %>%
  pivot_wider(names_from = away_covered, values_from = n)

merge(spread_home, spread_away, by.x="home_team_abbrev", by.y="away_team_abbrev") %>%
  mutate(cover_perc=(TRUE.x + TRUE.y)/(TRUE.x + TRUE.y + FALSE.x + FALSE.y)) %>%
  rename(team=home_team_abbrev) %>%
  arrange(desc(cover_perc)) %>%
  head(5) %>%
  select(team, cover_perc)
```
