---
title: "homework"
author: "Sean Li and Ben Thorpe"
date: "Due before" 
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The main goals of this homework is to let you practice data loading, data cleaning, and finding useful information. It will mimic the process you will go through in a real sports analysis endeavor. 

Remember, this is important because often times in the project, half the work is getting the right data! 


## Homework Instructions

Complete the exercises below through creating a new RMD file in the same repository you had your first homework done in. You are also welcome to create a new repository to do this homework too. 

Data for this homework can be found through the class repository. You can obtain it through downloading the file directly from the repository on Github or downloading the data from 
https://www.basketball-reference.com/leagues/NBA_2022_per_game.html


Once you are done, commit and push one last time to make sure all your changes are tracked. Then fill out the google form below. 

TURN IT IN HERE: 



## Exercises

```{r libraries}
library(tidyverse)
library(openxlsx)
```


### Loading in the Data

Today we will be working with 2021-2022 NBA Player Stats Data. 

1. Load in the data (make sure datafile is in repository folder!!)

2. Take a look at the data using glimpse()


### Rumble. 

2. How many players are in this dataset? (HINT: its not just the number of rows)

3. 








